---
title: "Major assignment"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
library(factoextra)
library(vegan)
library(chemometrics)
library(ggplot2)
library(cluster)
```
```{r}
Data<-read.csv("PlantData.csv")
PlotAge<-read.csv("PlotAge.csv")
```

Perform Hellinger transformation to give low weight to species with many zeros
```{r}

#Make plot names the row names
row.names(Data)=Data[,1]; Data[,1]<-NULL


#Hellinger transformation as some sites have many zeros
Data2<-decostand(Data,"hell", na.rm=TRUE)
```

Run the PCA and plot it. Divide plots by age class
```{r}
fit <- prcomp(Data2)

fviz_pca_biplot(fit,habillage=PlotAge$Age.class, label = "var")
```

```{r}
#Let's look at the amount of variation explained by each axis
summary(fit)
```

Lets test our assumptions

```{r}
#multivariate normality
#library("mvnormtest")
#mshapiro.test(Data2[,1:21])

#How many components to keep?
screeplot(fit)

#How much variance is explained by each species?
pcaVarexpl(X=Data2, a=2)
names(Data2)


#Check for outliers
fit2<-princomp(Data2)
pcaDiagplot(Data2,fit2,a=4)
```

Now let's do a PCoA. First, we need to contruct our axes

```{r}
#Set the seed to stop the values from changing 
set.seed(74)
#Create dissimilarity matrix using Manhattan distance
Matrix<-vegdist(Data2, "manhattan")

#Correct negative Eigenvalues
PCoA <- wcmdscale(Matrix, eig = TRUE,  add= TRUE) 

#Extract Eigenvalues
ev<-eigenvals(PCoA)
#Determine variation explained by each axis
AxisPer<-cumsum(ev/sum(ev))
#Extract site scores
points<-PCoA$points[,1:2]
#Extract species scores
sppco <- cor(Data2, PCoA$points)
#View percet of variation explaine dby each axis
AxisPer
#Extract species scores for first two axes
sppco[,1:2] 
#Create data frame of species scores
Variables<-data.frame(sppco) 
```

Now  lets plot the PCoA Ordination of species scores!
```{r}
#Add the first two principle components as columns to my dataset
Data2$PC1<-points[,1]
Data2$PC2<-points[,2]
Data2<-data.frame(Data2)

#Plot it!
PCoAPlot1<- ggplot(data=Data2,aes(x=PC1,y=PC2)) +
  geom_point(data=Data2,aes(x=PC1,y=PC2, shape=PlotAge[,2], col=PlotAge[,2]), size=3)+
   labs(x="PCoA Axis 1 (17.8%)", y = "PCoA Axis 2 (13.8%)")
PCoAPlot1
```


Now I want to find out how the species scores are influencing the sites. To do this, I will plot the species scores agaisnt the axes.
```{r}
PCoAPlot2<-ggplot() + 
  geom_text(data=Variables, aes(x=Dim1,y=Dim2,label=row.names(Variables),
                           hjust=0.5*(1-sign(Dim1)),vjust=0.5*(1-sign(Dim2))), size=4)

PCoAPlot2
```


test for multivariate normality
```{r}
apply(Data2, 2 ,shapiro.test)
```

